Squint: Lossy Hierarchical Compression on Symbol Streams

### 2012-12-10 tex/paper.tex
finished similarity vs cluster and reduction rule analysis, language needs polishing

@@

\begin{table}[t]
\subfloat[$\epsilon = .4$]{
\begin{tabular}{l|l|l|l|l||l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
\hline
rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
\end{tabular}
}

@@

Runtime Comparison, esp. w.r.t num of rules.
Quality Comparison.
Epsilon Comparison.

@@

\begin{table}[t]
\subfloat[$\epsilon = .4$]{
\begin{tabular}{l|l|l|l||l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
\end{tabular}
}

@@

Thus, we see that similarity provides no consistent advantage over cluster
and has much less scalable performance, with an estimated running time of 
BLAH. For the rest of the paper we will only test Cluster.

\subsection{Grammar Reduction Impact}
Although the grammars generated by Sequitur are irreducible,
our lossifiers produce grammars which could potentially be
compressed further (losslessly) using Reduction Rules without
changing the output string. To test the cost and
effectiveness of applying reduction rules, we looked at the
rep-n and gen-n test cases again, this time focusing on the
clustering algorithm for $\epsilon = .2$.

\begin{table}[t]
\subfloat[rep-n tests]{
\begin{tabular}{l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
\hline
rep-100   & & /& & /\\
rep-400   & & /& & /\\
rep-1600  & & /& & /\\
rep-6400  & & /& & /\\
rep-12800 & & /& & /\\
\end{tabular}
}

\subfloat[gen-n tests]{
\begin{tabular}{l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
\hline
gen-100   & & /& & /\\
gen-400   & & /& & /\\
gen-1600  & & /& & /\\
gen-6400  & & /& & /\\
gen-12800 & & /& & /\\
\end{tabular}
}
\caption{Reducer Tests for $\epsilon = .2$}
\end{table}

>>>>

\subsubsection{Rep-n results}

@@

\begin{table}[p]
\subfloat[$\epsilon = .4$]{
\begin{tabular}{l|l|l|l|l||l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & S-P(roportion) & C-T & C-R & C-L & C-P \\
\hline
rep-100   & 0.006 & 18/34    & 0.713 & 0.07 & 0.002 & 20/34    & 0.535 & 0.04\\
rep-400   & 0.028 & 45/76    & 0.269 & 0.01 & 0.009 & 41/76    & 0.743 & 0.04\\
rep-1600  & 0.355 & 103/194  & 0.873 & 0.03 & 0.042 & 103/194  & 0.804 & 0.02\\
rep-6400  & 3.978 & 258/448  & 0.706 & 0.005& 0.203 & 256/448  & 0.869 & 0.50\\
rep-12800 & 44.63 & 604/1064 & 0.791 & 0.01 & 1.136 & 600/1064 & 0.856 & 0.01\\
rep-25600 & 135.0 & 900/1482 & 0.829 & 0.005& 2.906 & 898/1482 & 0.725 & 0.25\\
\end{tabular}
}

@@

The compression ratios achieved for both algorithms are roughly equal for the same
value for $\epsilon$, so it is fair to compare the algorithms for the same $\epsilon$.

From plotting the data, we see that runtime of the algorithms are roughly polynomial
in the size of the data, so we can perform a linear regression on the log-log transform
of the data to find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
length of the input string. In the chart below, all of the linear regressions had
$R^2$ values of over $0.98$.

\begin{tabular}{l|l|l|l|l}
Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent & $1.859$ & $2.026$ & $1.306$ & $1.305$
\end{tabular}

For both values of $\epsilon$, 
The clustering algorithm is nearly linear, while the similarity algorithm is roughly
quadratic. On this dataset, the leveshtein distances were similar for both algorithms
but fairly high since the length of the final string could change drastically.
The proportion of '*'s in the input was usually the right order of magnitude,
with both algorithms making big mistakes occassionally. In most cases,
the two algorithms maintained similarly acceptable fidelity levels in recreating
the texture of the original rep-n input file.

Setting higher values for $\epsilon$ had diminishing returns for compression
for larger and larger file sizes, at the cost of drastically lowering the fidelity
of the output.

\subsubsection{Gen-n results}

@@

\begin{table}[p]
\subfloat[$\epsilon = .4$]{
\begin{tabular}{l|l|l|l||l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
gen-100   & 0.006 & 87/89    & 0.04 &0.006 & 87/89    & 0.04\\
gen-400   & 0.175 & 224/258  & 0.12 &0.069 & 219/258  & 0.19\\
gen-1600  & 4.863 & 539/697  & 0.31 &0.732 & 529/697  & 0.33\\
gen-6400  & 216.9 & 1740/2348& 0.41 &9.127 & 1694/2348& 0.43\\
gen-12800 & 1812  & 3296/4523& 0.40 &29.17 & 3181/4523& 0.44\\
gen-25600 & 12989 & 5823/8060& 0.43 &176.7 & 5540/8060& 0.50\\
\end{tabular}
}

@@

Just as with the rep-n data, log-log regression tests yield the following
exponents for the runtime on the gen-n data.

\begin{tabular}{l|l|l|l|l}
Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent & $2.629$ & $2.493$ & $1.814$ & $2.061$
\end{tabular}

Again, the two algorithms have almost identical compression ratios and
their fidelity is about the same both subjectively and based on levenshtein
distance.

INCLUDE SAMPLE TEXT

The runtimes are noticeably worse than for the rep-n dataset. The
similarity algorithm runs in roughly $n^{2.5}$ time while cluster
runs in roughly $n^2$ time.

Changing epsilon in this case had a large effect on both the compression
ratio and the fidelity of the final output string.

\subsubsection{Summary}

Both of the algorithms were able to achieve great additional compression
ratios on top of \Sequitur of up to about 40 percent for the highly
structured rep-n data, and 30 percent for the more loosely structured
gen-n data without using the grammar reducer. The fidelity between
the original and lossified strings wasn't great as measured by
levenshtein distance, but the texture in both the rep-n and
gen-n datasets was preserved well. Decreasing the $\epsilon$ threshold
improved fidelity greatly for larger very structured data at little
cost to the compression ratio, but hurt the compression ratio
a lot for less structured data.

The differences between the two algorithms lie almost entirely in their
runtime. Both algorithms perform much worse on the less structued
gen-n data, we believe because simplifications are made more incrementally
and the grammar does not get dramatically simpler after a few iterations
of our algorithm like it does for the rep-n data.
Unfortunately the Similarity algorithm tends to consistently incur
an overhead of about $O(\sqrt{n})$ on top of the runtime for the
Cluster algorithm. Because of this we dismiss the Similarity algorithm
as unscalable and do not test it further.

\subsection{Grammar Reduction Impact}
Although the grammars generated by Sequitur are irreducible,
our lossifiers produce grammars which could potentially be
compressed further (losslessly) using Reduction Rules without
changing the output string. To test the cost and
effectiveness of applying reduction rules, we looked at the
rep-n and gen-n test cases again, this time focusing on the
clustering algorithm for $\epsilon = .2$. This time we give the ratio
on top of the simplified grammar produced by the clustering algorithm,
and the time taken to run just the reduction rules.

\begin{table}[p]
\subfloat[rep-n tests]{
\begin{tabular}{l|l|l}
Dataset & Time & Ratio \\
\hline
rep-100   & 0.008 & 29/30  \\
rep-400   & 0.028 & 56/60  \\
rep-1600  & 0.132 & 96/123 \\
rep-6400  & 1.145 & 153/186\\
rep-12800 & 19.43 & 403/632\\
\end{tabular}
}
\subfloat[gen-n tests]{
\begin{tabular}{l|l|l}
Dataset & Time & Ratio \\
\hline
gen-100   & 0.016 & 89/89\\
gen-400   & 0.126 & 245/245\\
gen-1600  & 4.833 & 660/664\\
gen-6400  & 308.7 & 2139/2171\\
\end{tabular}
}
\caption{Reducer Tests for $\epsilon = .2$}
\end{table}

When plotted on a log-log scale, the runtime for our straightforward 
implementation of the reduction
rules appears to be superpolynomial. While modest additional gains of about 30 percent
are possible for highly self-similar data, very few additional gains are
possible for more diverse datasets such as gen-n. This appears to be because
the only simplifications that are possible on diverse datasets don't lead to
the wide-ranging repetitive substitutions which allow reduction rules to be allowed.
The simplifications in gen-n tend to be more local.

Because we were unable to get reduction rules to scale to larger datasets,
we will not consider them further.
